---
title: "fast-fires"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("setup.R")
```

Summary of Fast Fire impacts from the ICS-209-PLUS (1999-2020). 

First, we need to handle possible multiple entries due to the reporting nature for complex fires. If this is not done, there are likely double counts for attributes like burned area and destroyed structures.

```{r}
icsdir <- '/Users/max/Library/CloudStorage/OneDrive-Personal/mcook/ics209-plus-fired/data/'
# Load the complex associations table
complex.df <- suppressMessages(
 read_csv(paste0(
  icsdir,"tabular/raw/wf-incidents/ics209-plus_v2.0/ics209plus-wildfire/ics209-plus-wf_complex_associations_1999to2020.csv"
 ), show_col_types = FALSE)
)

# Isolate non-complex fires
no.complexes <- ics %>% 
 # Remove some known "problem fires"
 filter(INCIDENT_ID != "2017_7211255_REDWOOD VALLEY INCIDENT",
        INCIDENT_ID != "2017_7293073_REDWOOD VALLEY INCIDENT") %>%
 # Filter incidents with 0 acres
 filter(FINAL_ACRES!=0,
        # remove incidents found in the complex associations table
        ! INCIDENT_ID %in% complex.df$MEMBER_INCIDENT_ID,  # member incidents
        ! INCIDENT_ID %in% complex.df$CPLX_INCIDENT_ID,  # complex fires
        # filter to fires where FPA-FOD shows 1 event or NULL
        FOD_FIRE_NUM==1 | is.na(FOD_FIRE_NUM))

# Check the number of complex fires in the complex association table
length(unique(complex.df$CPLX_INCIDENT_ID))

# Isolate complex fires
complexes <- ics %>%
 filter(INCIDENT_ID %in% complex.df$CPLX_INCIDENT_ID)

# Check the number of complex fires in that match
dim(complexes)

# Bind back to the ICS209-PLUS
ics <- bind_rows(complexes,no.complexes)
rm(complexes,no.complexes,complex.df)
gc()
```

Summary statistics from the ICS209-PLUS:

```{r}
# Observed fast fires break and number of homes threatened, damaged, destroyed
ics.fast <- ics %>% 
 mutate(WF_MAX_FSR_HA = WF_MAX_FSR*0.404686) %>% 
 filter(WF_MAX_FSR_HA >= 1620) %>%
 filter(START_YEAR >= 2000)

# Summarize number of threatened, damaged, destroyed structures by fast fires from the 209s
print("ICS-209-PLUS 1999-2020 Fast Fires, Structural Impacts: ")
paste0("Total fast fires (N): ",dim(ics.fast)[1])
print("~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~")
paste0("Total Structures Threatened: ", scales::comma(sum(ics.fast$STR_THREATENED_MAX, na.rm=TRUE)))
paste0("Total Structures Damaged: ", scales::comma(sum(ics.fast$STR_DAMAGED_TOTAL, na.rm=TRUE)))
paste0("Total Structures Destroyed: ", scales::comma(sum(ics.fast$STR_DESTROYED_TOTAL, na.rm=TRUE)))
paste0("Total Residential Structures Destroyed: ",scales::comma(sum(ics.fast$STR_DESTROYED_RES_TOTAL, na.rm=TRUE)))
print("~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~")
# Percent of homes threatened, damaged, destroyed
paste0("Percent of Structures Threatened: ",
      sum(ics.fast$STR_THREATENED_MAX, na.rm=T)/sum(ics$STR_THREATENED_MAX, na.rm=T)*100)
paste0("Percent of Structures Damaged: ", 
      sum(ics.fast$STR_DAMAGED_TOTAL, na.rm=T)/sum(ics$STR_DAMAGED_TOTAL, na.rm=T)*100)
paste0("Percent of Structures Destroyed: ", 
      sum(ics.fast$STR_DESTROYED_TOTAL, na.rm=T)/sum(ics$STR_DESTROYED_TOTAL, na.rm=T)*100)
paste0("Percent of Residential Structures Destroyed: ", 
      sum(ics.fast$STR_DESTROYED_RES_TOTAL, na.rm=T)/sum(ics$STR_DESTROYED_RES_TOTAL, na.rm=T)*100)
print("~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~")
paste0("Total Suppression Costs: ",scales::comma(sum(ics.fast$PROJECTED_FINAL_IM_COST, na.rm=T)))
paste0("Percent of Suppression Costs: ",
      scales::comma(sum(ics.fast$PROJECTED_FINAL_IM_COST, na.rm=T)/
                     sum(ics$PROJECTED_FINAL_IM_COST, na.rm=T)*100))
print("~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~")
paste0("Total Fatalities:",sum(ics.fast$FATALITIES,na.rm=T))
paste0("Percent of Fatalities:",sum(ics.fast$FATALITIES,na.rm=T)/sum(ics$FATALITIES,na.rm=T)*100)
paste0("Total Injuries: ",sum(ics.fast$INJURIES_TOTAL,na.rm=T))
paste0("Percent of Injuries: ",sum(ics.fast$INJURIES_TOTAL,na.rm=T)/sum(ics$INJURIES_TOTAL,na.rm=T)*100)
# Write out
st_write(ics.fast, "../../data/spatial/mod/ics209plus_conus_fast_2000to2020.gpkg", delete_dsn=T)
```

Subset FIRED events to fast fires (max daily growth > 1620ha). Exclude cropland systems.

```{r include=F}
# Load the BUPR summaries
bupr <- read_csv("../../data/tabular/bupr_sums.csv",show_col_types = FALSE) %>%
 dplyr::select(id,bupr_sum,bupr_sum1km,bupr_sum4km)

# Remove fired that occurred predominantly in cropland landcover
fired.fast <- fired %>% 
 dplyr::select(-c(bupr_sum,bupr_sum1k)) %>%
 filter(lc_name != "Croplands") %>%
 mutate(tot_ar_ha = tot_ar_km2*100,
        mx_grw_ha = mx_grw_km2*100,
        mx_grw_pct = (mx_grw_ha / tot_ar_ha)*100) %>% 
 filter(mx_grw_ha >= 1620) %>%
 st_transform(st_crs(lambert.prj)) %>%
 # Join to BUPR summary
 left_join(bupr, by="id")

# Get the top 100 fastest fires in FIRED
fired.t100 <- top_n(fired.fast, 100, mx_grw_ha)

# Write out the top 100 fires
st_write(fired.fast, "../../data/spatial/mod/conus_fast-fires_2001to2020.gpkg", delete_dsn=T)
st_write(fired.t100, "../../data/spatial/mod/fastest100_fired.gpkg", delete_dsn=T)
```

Set up the Table 1 and Supplemental Table 3:

Addin the East Amarillo Complex

```{r}
# Load the QC FIRED (matches the ICS-FIRED version):
fired.fast.qc <- st_read("../../data/spatial/mod/conus_fast-fires_qc_2001to2020.gpkg") %>%
 rename(FIRED_ID = id) %>%
 st_set_geometry(NULL) %>% as_tibble()

# Add in the EAST AMARILLO COMPLEX fire manually
addin.fired <- fired.fast.qc %>% filter(FIRED_ID == 40312) %>%
 mutate(INCIDENT_ID = "2006_TX-TXS-066077_EAST AMARILLO COMPLEX")

addin <- ics %>%
 st_set_geometry(NULL) %>% as_tibble() %>%
 filter(INCIDENT_ID == "2006_TX-TXS-066077_EAST AMARILLO COMPLEX") %>%
 left_join(addin.fired, by="INCIDENT_ID") %>%
 select(
  -c(bupr_sum,bupr_sum1km,bupr_sum4km,lc_name,lc_description,GACCAbbrev,
     NA_L3CODE,NA_L3NAME,NEON_DOMAIN,x,y,tot_ar_ha,mx_grw_ha,mx_grw_pct)) %>%
 mutate(ig_date = as.Date(ig_date),
        last_date = as.Date(last_date),
        mx_grw_dte = as.Date(mx_grw_dte))

# Bind back to the ICS-FIRED
ics.fired <- ics.fired %>% bind_rows(addin)

rm(addin.fired,addin,fired.fast.qc)
```


```{r include=F}
# Load the QC FIRED (matches the ICS-FIRED version):
fired.fast.qc <- st_read("../../data/spatial/mod/conus_fast-fires_qc_2001to2020.gpkg") %>%
 rename(FIRED_ID = id) %>%
 select(FIRED_ID,bupr_sum,bupr_sum1km,bupr_sum4km,lc_name,lc_description) %>%
 st_set_geometry(NULL) %>% as_tibble()

# Get the top 100 fastest fires in ICS-FIRED
ics.fired.t100 <- ics.fired %>%
 filter(STUSPS != 'AK') %>%
 # WITCH is in the new WITCH-POOMACHA record
 filter(INCIDENT_NAME != "WITCH") %>% 
 mutate(mx_grw_ha = mx_grw_km2*100,
        tot_ar_ha = tot_ar_km2*100,
        mx_grw_pct = (mx_grw_ha / tot_ar_ha)*100) %>%
 filter(mx_grw_ha >= 1620) %>%
 top_n(., 100, mx_grw_ha) %>%
 st_transform(st_crs(lambert.prj))
# Write to GeoPackage
st_write(ics.fired.t100, "../../data/spatial/mod/fastest100_ics-fired.gpkg", delete_dsn=T)

# Set up the cleaned table, attaching BUPR and MCD12Q1
ics.fired.t100 <- ics.fired.t100 %>%
 left_join(., fired.fast.qc, by="FIRED_ID") %>%
 # Manually fix some missing data (unsure why)
 mutate(lc_name = if_else(INCIDENT_NAME == "LNU LIGHTNING COMPLEX", "Grasslands", lc_name),
        bupr_sum4km = if_else(INCIDENT_NAME == "LNU LIGHTNING COMPLEX", 34344, bupr_sum4km)) %>%
 # Convert all fire names to capitol letters
 mutate(INCIDENT_NAME = toupper(INCIDENT_NAME))

# Clean up the table
cleaned.t100 <- ics.fired.t100 %>%
 st_set_geometry(NULL) %>% as_tibble() %>%
 mutate(FINAL_HA = FINAL_ACRES*0.404686) %>%
 select(INCIDENT_NAME,ig_year,STUSPS,FINAL_HA,mx_grw_ha,event_dur,PROJECTED_FINAL_IM_COST,bupr_sum4km,STR_DESTROYED_TOTAL,
        TOTAL_AERIAL_SUM,TOTAL_PERSONNEL_SUM,lc_name) 

# Write out the top 100 to a CSV
rm(fired.fast.qc)
```

Check how many fires are not matched between FIRED and ICS-FIRED:

```{r}
fired.fast.qc <- st_read("../../data/spatial/mod/conus_fast-fires_qc_2001to2020.gpkg")
dim(fired.fast.qc %>% filter(id %in% ics.fired.t100$FIRED_ID))
rm(fired.fast.qc)
```

Get some statistics for the table captions:

```{r}
paste0("Suppression costs for top 100: ",prettyNum(sum(ics.fired.t100$PROJECTED_FINAL_IM_COST,na.rm=T),big.mark=",",scientific=F))
paste0("Exposed properties (within 4km): ",prettyNum(sum(ics.fired.t100$bupr_sum4km,na.rm=T),big.mark=",",scientific=F))
paste0("Destroyed structures: ",prettyNum(sum(ics.fired.t100$STR_DESTROYED_TOTAL,na.rm=T),big.mark=",",scientific=F))
summary(factor(ics.fired.t100$lc_name))
```

Table 1: Top 20 fastest fires (CONUS)

Set up the table for export in Science formatting:

```{r}

# Get the top 20
cleaned.t20 <- cleaned.t100 %>%
 top_n(., 20, mx_grw_ha)

# Set font name for table
fontname <- "Times New Roman"

# Clean up extra rows and digits
cleaned.t1 <- cleaned.t20 %>%
 mutate(ig_year = as.character(ig_year),
        PROJECTED_FINAL_IM_COST = if_else(PROJECTED_FINAL_IM_COST < 100, 0, PROJECTED_FINAL_IM_COST)) %>%
 mutate_if(is.double, ~ round(., digits = 0)) %>%
 # Sort from fastest to slowest
 arrange(desc(mx_grw_ha))

# Fix names and add units
names(cleaned.t1) <- c("Incident Name",
                        "Year",
                        "State",
                        "Fire Size (ha)",
                        "Maximum Fire Growth (ha / day)",
                        "Fire Duration (days)",
                        "Cost ($)",
                        "Properties Exposed (within 4km)",
                        "Structures Destroyed",
                        "Total Aerial Units",
                        "Total Personnel",
                        "Dominant Vegetation")

# Create the flextable
ft1 <- flextable(cleaned.t1) %>%
 font(fontname = fontname, part = "all") %>%
 autofit() %>% fit_to_width(7.5)
ft1
print(ft1, preview = "docx")
# Write to a CSV
write_csv(cleaned.t1, "../../data/tabular/ics-fired_t20_cleaned.csv")

```

Supplemental Table S3: Top 100 fastest fires (CONUS)

Set up the table for export in Science formatting:

```{r}

# Set font name for table
fontname <- "Times New Roman"

# Clean up extra rows and digits
cleaned.ts3 <- cleaned.t100 %>%
 mutate(ig_year = as.character(ig_year),
        PROJECTED_FINAL_IM_COST = if_else(PROJECTED_FINAL_IM_COST < 100, 0, PROJECTED_FINAL_IM_COST)) %>%
 mutate_if(is.double, ~ round(., digits = 0)) %>%
 # Sort from fastest to slowest
 arrange(desc(mx_grw_ha))

# Fix names and add units
names(cleaned.ts3) <- c("Incident Name",
                        "Year",
                        "State",
                        "Fire Size (ha)",
                        "Maximum Fire Growth (ha / day)",
                        "Fire Duration (days)",
                        "Cost ($)",
                        "Properties Exposed (within 4km)",
                        "Structures Destroyed",
                        "Total Aerial Units",
                        "Total Personnel",
                        "Dominant Vegetation")

# Create the flextable
fts3 <- flextable(cleaned.ts3) %>%
  font(fontname = fontname, part = "all")
fts3
print(fts3, preview = "docx")
# Write to a CSV
write_csv(cleaned.ts3, "../../data/tabular/ics-fired_t100_cleaned.csv")

```

```{r}
# Print some statistics of the fast fires subset
print("1.7% of fires in the FIRED database meet the fast fires threshold ...")
1620/95888*100
print("~~~~~~~~~~~Summary of Max Growth~~~~~~~~~~~~~")
summary(fired.fast$mx_grw_km2)
print("~~~~~~~~~~~BURNED AREA~~~~~~~~~~~~~")
paste0("% of total burned area: ",
      sum(fired.fast$tot_ar_km2)/sum(fired$tot_ar_km2)*100)
print("~~~~~~~~~~~BUPR~~~~~~~~~~~~~")
paste0("Number of exposed properties within perimeter: ", sum(fired.fast$bupr_sum,na.rm=T))
paste0("Number of exposed properties within 1km: ", sum(fired.fast$bupr_sum1k,na.rm=T))
paste0("Number of exposed properties within 4km: ", sum(fired.fast$bupr_sum4k,na.rm=T))
```

Figure 1. Spatial distribution of fast fires across CONUS. Top 100 fastest fires are highlighted by size/color.

```{r warning=F, message=F}

# Bring in cities and lakes
cities <- st_read(paste0(maindir,"/data/boundaries/political/USGS/USA_MajorCities.shp")) %>%
 st_transform(st_crs(lambert.prj)) %>%
 filter(NAME == "Denver")
lakes <- st_read(paste0(maindir,"/data/water/great_lakes.gpkg"))

# Spatial map of top 100
centroid1 <- st_centroid(fired.fast)
centroid2 <- st_centroid(fired.t100)
centroid3 <- st_centroid(ics.fired.t100)

f1 <- ggplot() + 
   geom_sf(data = states, lwd=0.7, color=alpha("gray20", 0.85), fill=alpha("white",0.75)) +
   geom_sf(data = states %>% st_buffer(1e5), color=NA, fill=NA) +
   geom_sf(data = centroid1, color=alpha("gray40", 1), size=0.3) +
   geom_sf(data = centroid2, aes(color=mx_grw_ha, size=mx_grw_ha)) +
   # geom_sf(data=cities, size=2.5, shape=18) +
   # geom_sf(data=lakes, color=NA, fill="grey75") +
   scale_color_viridis_c(trans="log", option="inferno", direction=-1,
                         breaks=c(20000,40000,80000,160000),
                         labels=c("20k","40k","80k","160k")) +
   labs(fill="", color="Maximum Single-Day Growth (ha/day)", tag=expression(bold("(A)"))) +
   guides(color = guide_colourbar(direction = "horizontal", barwidth = 12, barheight = 0.60, 
                                  ticks=F, title.position = "top"),
          label.theme = element_text(angle = 0, size = 9),
          size="none") +
   theme_void() +
   theme(plot.margin = unit(c(0.0,0.0,0.0,0.0), "in"),
         plot.title = element_blank(),
         legend.title = element_text(angle = 0, size=10),
         legend.text = element_text(size=9),
         legend.position=c(0.23, 0.08),
         plot.tag = element_text(vjust = -1.2),
         text = element_text(family = "Helvetica")) +
   ggspatial::annotation_scale(location = "br", width_hint = 0.1,
                               pad_x = unit(0.2,"in"), pad_y = unit(0.2,"in"),
                               line_width = 0.5, text_pad = unit(0.15,"cm"),
                               height = unit(0.15,"cm")) +
   ggspatial::annotation_north_arrow(location = "br", which_north = "true",
                                     pad_x = unit(0.3,"in"), pad_y= unit(0.4,"in"),
                                     width = unit(0.8,"cm"), height = unit(0.8,"cm"),
                                     style = north_arrow_fancy_orienteering)
f1

# Save it out
ggsave(f1, file = "../../figures/Figure1A_fastest_fires_conus.png",
       dpi = 300, bg="white") # adjust dpi accordingly

rm(f1,centroid1,centroid2,centroid3,cities,lakes)
```

Figure 1B-D. Case studies - selection of fast fires from the top 100 with ZTRAX point locations and daily burned area.

Load the ZTRAX point data and the list of fire boundaries (using MTBS):

```{r}
# ZTRAX point locations
# ztrax <- st_read("../../data/spatial/mod/ztrax/extract/zasmt_fast_fires_extract.shp") %>%
#  st_transform(st_crs(mtbs))
# # Selected MTBS fires
# mtbs_ <- mtbs %>%
#  filter(Event_ID %in% c(
#   # NW Oklahoma Complex, OK & KS (2017)
#   "OK3671009983720170307","OK3677410017520170307",
#   # Cold Springs, WA (2020)
#   "WA4836011949320200907",
#   # Martin Fire, NV (2018)
#   "NV4152811742720180705",
#   # Cedar Fire, CA (2003)
#   "CA3293911676620031025"))
```

Load the HISDAC-US BUPR points:

```{r}
hisdac <- st_read("../../data/spatial/mod/hisdac/bupr_random_points_case_study.gpkg") %>% st_transform(st_crs(lambert.prj))
```

NW Oklahoma Complex, OK/KS (2017):

```{r}

# Grab the start and end DOY
evt <- fired.fast %>% 
 filter(id %in% c(33409,33411)) %>%
 mutate(last_day = lubridate::yday(last_date))
start_doy = min(evt$ig_day) # first DOY
end_doy = max(evt$last_day) # end DOY

# Set the geometries
st_agr(mtbs) <- "constant"
st_agr(hisdac) <- "constant"

# MTBS footprint
fire <- mtbs %>%
  filter(Event_ID %in% c("OK3671009983720170307","OK3677410017520170307")) %>%
  st_transform(st_crs(lambert.prj))

# Intersect the HISDAC
points <- hisdac %>% st_intersection(.,fire)

# Convert to SpatVects
fire <- vect(fire)
points <- vect(points)

# Load the burndate raster (MCD64A1)
# Reproject the grid and crop to the fire perimeter
fire_dir <- paste0(maindir,"/fast-fires/data/spatial/mod/mcd64a1/")
mcd64 <- terra::rast('../../data/spatial/mod/mcd64a1/mcd64a1_annual_burndate_y2017.tif')
mcd64 <- terra::project(mcd64,crs(lambert.prj, proj=TRUE))
mcd64 <- terra::crop(mcd64,fire,mask=T)

# Convert to data frame
p1 <- mcd64 %>% as.data.frame(xy = TRUE) %>% mutate(
 value = as.integer(mcd64a1_annual_burndate_y2017)
) %>% 
 dplyr::select(-mcd64a1_annual_burndate_y2017) %>%
 mutate(value = if_else(value < start_doy, NA, value),
        value = if_else(value > end_doy, NA, value),
        value = value - start_doy,
        value = if_else(value > 5, 6, value)) %>%
 filter(value > 0) %>%
 drop_na() %>%
 ggplot() +
 geom_raster(aes(x = x, y = y, fill = value)) +
 # scale_fill_brewer(palette="Reds", na.value="white", direction=-1) +
 scale_fill_fermenter(breaks = c(1:5,6), palette = "Reds") + 
 # Add the fire boundary
 geom_sf(data=fire, fill="transparent", color="gray30", lwd=0.45) +
 geom_sf(data=points, color=alpha("black",0.8), size=0.85, shape=18) +
 coord_sf() +
 # , tag=expression(bold("(B)"))
 labs(fill="Days Since Ignition") +
 theme_void() +
 theme(legend.position="none",
       plot.title = element_text(hjust = 0.5),
       plot.tag = element_text(hjust = -2.5),
       text = element_text(family = "Helvetica")) +
 guides(fill = element_blank()) +
 # guides(fill = guide_colorbar(title.position="top", title.hjust = 0.5,
 #                              barwidth = 12, barheight = 0.5,
 #                              ticks=FALSE, draw.llim=FALSE)) +
 ggspatial::annotation_scale(location = "br", width_hint = 0.1,
                             pad_x = unit(0.2,"in"), pad_y = unit(0.2,"in"),
                             line_width = 0.5, text_pad = unit(0.15,"cm"),
                             height = unit(0.15,"cm")) +
 ggspatial::annotation_north_arrow(location = "br", which_north = "true",
                                   pad_x = unit(0.3,"in"), pad_y= unit(0.4,"in"),
                                   width = unit(0.8,"cm"), height = unit(0.8,"cm"),
                                   style = north_arrow_fancy_orienteering)
p1
gc()

# Save the image
ggsave(p1, file = "../../figures/FigureX_NWOKCMPLXFire.png",
       width=3.5, height=3.5, dpi = 300, bg="transparent") # adjust dpi accordingly

```

Cold Springs Fire, WA (2020)

```{r}

# Grab the start and end DOY
start_doy = 250 # first DOY
end_doy = 259 # end DOY

# Set the geometries
st_agr(mtbs) <- "constant"
st_agr(hisdac) <- "constant"

# MTBS footprint
fire <- mtbs %>%
  filter(Event_ID %in% c("WA4836011949320200907")) %>%
  st_transform(st_crs(lambert.prj))

# Intersect the HISDAC
points <- hisdac %>% st_intersection(.,fire)

# Convert to SpatVects
fire <- vect(fire)
points <- vect(points)

# Load the burndate raster (MCD64A1)
# Reproject the grid and crop to the fire perimeter
fire_dir <- paste0(maindir,"/fast-fires/data/spatial/mod/mcd64a1/")
mcd64 <- terra::rast('../../data/spatial/mod/mcd64a1/mcd64a1_annual_burndate_y2020.tif')
mcd64 <- terra::project(mcd64,crs(lambert.prj, proj=TRUE))
mcd64 <- terra::crop(mcd64,fire,mask=T)

# Convert to data frame
p2 <- mcd64 %>% as.data.frame(xy = TRUE) %>% mutate(
 value = as.integer(mcd64a1_annual_burndate_y2020)
) %>% 
 dplyr::select(-mcd64a1_annual_burndate_y2020) %>%
 mutate(value = if_else(value < start_doy, NA, value),
        value = if_else(value > end_doy, NA, value),
        value = value - start_doy,
        value = if_else(value > 5, 6, value)) %>%
 drop_na() %>%
 ggplot() +
 geom_raster(aes(x = x, y = y, fill = value)) +
 # scale_fill_brewer(palette="Reds", na.value="white", direction=-1) +
 scale_fill_fermenter(breaks = c(1:5,6), palette = "Reds") +
 # Add the fire boundary
 geom_sf(data=fire, fill="transparent", color="gray30", lwd=0.45) +
 geom_sf(data=points, color=alpha("black",0.8), size=0.85, shape=18) +
 coord_sf() +
 # , tag=expression(bold("C"))
 labs(fill="Days Since Ignition") +
 theme_void() +
 theme(legend.position="none",
       plot.title = element_text(hjust = 0.5),
       plot.tag = element_text(hjust = 1.2),
       text = element_text(family = "Helvetica")) +
 # guides(fill = guide_colorbar(title.position="top", title.hjust = 0.5,
 #                               barwidth = 12, barheight = 0.5,
 #                               ticks=FALSE, draw.llim=FALSE)) +
 ggspatial::annotation_scale(location = "br", width_hint = 0.1,
                             pad_x = unit(0.0,"in"), pad_y = unit(0.2,"in"),
                             line_width = 0.5, text_pad = unit(0.15,"cm"),
                             height = unit(0.15,"cm")) +
 ggspatial::annotation_north_arrow(location = "br", which_north = "true",
                                   pad_x = unit(0.1,"in"), pad_y= unit(0.4,"in"),
                                   width = unit(0.8,"cm"), height = unit(0.8,"cm"),
                                   style = north_arrow_fancy_orienteering)
p2
gc()

# Save the image
ggsave(p2, file = "../../figures/FigureX_ColdSpringsFire.png",
       width=3.5, height=3.5, dpi = 300, bg="transparent") # adjust dpi accordingly

```

Witch-Poomacha Fire, CA (2007) - highly destructive event

```{r}

# Grab the start and end DOY
evt <- fired.fast %>% 
 filter(id %in% c(46167)) %>%
 mutate(last_day = lubridate::yday(last_date))
start_doy = 291 # first DOY
end_doy = max(evt$last_day) # end DOY

# Set the geometries
st_agr(mtbs) <- "constant"
st_agr(hisdac) <- "constant"

# MTBS footprint
fire <- mtbs %>%
  filter(Event_ID %in% c("CA3307911676620071021","CA3330011692020071023")) %>%
  st_transform(st_crs(lambert.prj))

# Intersect the HISDAC
points <- hisdac %>% st_intersection(.,fire)

# Convert to SpatVects
fire <- vect(fire)
points <- vect(points)

# Load the burndate raster (MCD64A1)
# Reproject the grid and crop to the fire perimeter
fire_dir <- paste0(maindir,"/fast-fires/data/spatial/mod/mcd64a1/")
mcd64 <- terra::rast('../../data/spatial/mod/mcd64a1/mcd64a1_annual_burndate_y2007.tif')
mcd64 <- terra::project(mcd64,crs(lambert.prj, proj=TRUE))
mcd64 <- terra::crop(mcd64,fire,mask=T)

# Convert to data frame
p3 <- mcd64 %>% as.data.frame(xy = TRUE) %>% mutate(
 value = as.integer(mcd64a1_annual_burndate_y2007)
) %>% 
 dplyr::select(-mcd64a1_annual_burndate_y2007) %>%
 mutate(value = if_else(value < start_doy, NA, value),
        value = if_else(value > end_doy, NA, value),
        value = value - start_doy,
        value = if_else(value > 5, 6, value)) %>%
 filter(value > 0) %>%
 drop_na() %>%
 ggplot() +
 geom_raster(aes(x = x, y = y, fill = value)) +
 # scale_fill_brewer(palette="Reds", na.value="white", direction=-1) +
 scale_fill_fermenter(breaks = c(1:5,6), palette = "Reds") + 
 # Add the fire boundary
 geom_sf(data=fire, fill="transparent", color="gray30", lwd=0.45) +
 geom_sf(data=points, color=alpha("black",0.45), size=0.65, shape=18) +
 coord_sf() +
 # , tag=expression(bold("D"))
 labs(fill="Days Since Ignition") +
 theme_void() +
 theme(legend.position="none",
       plot.title = element_text(hjust = 0.5),
       plot.margin = unit(c(1,1,1,1),'lines'),
       plot.tag = element_text(hjust = 1.2),
       text = element_text(family = "Helvetica")) +
 # guides(fill = guide_colorbar(title.position="top", title.hjust = 0.5,
 #                               barwidth = 12, barheight = 0.5,
 #                               ticks=FALSE, draw.llim=FALSE)) +
 ggspatial::annotation_scale(location = "br", width_hint = 0.1,
                             pad_x = unit(0.1,"in"), pad_y = unit(0.15,"in"),
                             line_width = 0.5, text_pad = unit(0.15,"cm"),
                             height = unit(0.15,"cm")) +
 ggspatial::annotation_north_arrow(location = "br", which_north = "true",
                                   pad_x = unit(0.1,"in"), pad_y= unit(0.25,"in"),
                                   width = unit(0.8,"cm"), height = unit(0.8,"cm"),
                                   style = north_arrow_fancy_orienteering)
p3
gc()

# Save the image
ggsave(p3, file = "../../figures/FigureX_WitchPoomachaFire.png",
       dpi = 300, bg="transparent") # adjust dpi accordingly

```

3-Panel combined:

```{r}
ggpubr::ggarrange(p1, p2, p3, nrow = 1, common.legend = T, widths = c(2, 2, 2))
```

Additional fires:

Martin Fire, NV (2018)

```{r}

# Grab the start and end DOY
evt <- fired.fast %>% 
 filter(id %in% c(122367)) %>%
 mutate(last_day = lubridate::yday(last_date))
start_doy = min(evt$ig_day) # first DOY
end_doy = max(evt$last_day) # end DOY

# Set the geometries
st_agr(mtbs) <- "constant"
st_agr(ztrax) <- "constant"

# MTBS footprint
fire <- mtbs %>%
  filter(Event_ID %in% c("NV4152811742720180705")) %>%
  st_transform(st_crs(lambert.prj))

# ZTRAX point locations
points <- ztrax %>% 
  filter(YearBuilt <= 2018) %>%
  st_intersection(., fire) %>%
  st_transform(st_crs(lambert.prj))

# Convert to SpatVects
fire <- vect(fire)
points <- vect(points)

# Load the burndate raster (MCD64A1)
# Reproject the grid and crop to the fire perimeter
fire_dir <- paste0(maindir,"/fast-fires/data/spatial/mod/mcd64a1/")
mcd64 <- terra::rast('../../data/spatial/mod/mcd64a1/mcd64a1_annual_burndate_y2018.tif')
mcd64 <- terra::project(mcd64,crs(lambert.prj, proj=TRUE))
mcd64 <- terra::crop(mcd64,fire,mask=T)

# Convert to data frame
p3 <- mcd64 %>% as.data.frame(xy = TRUE) %>% mutate(
 value = as.integer(mcd64a1_annual_burndate_y2018)
) %>% 
 dplyr::select(-mcd64a1_annual_burndate_y2018) %>%
 mutate(value = if_else(value < start_doy, NA, value),
        value = if_else(value > end_doy, NA, value),
        value = value - start_doy,
        value = if_else(value > 5, 6, value)) %>%
 filter(value > 0) %>%
 drop_na() %>%
 ggplot() +
 geom_raster(aes(x = x, y = y, fill = value)) +
 # scale_fill_brewer(palette="Reds", na.value="white", direction=-1) +
 scale_fill_fermenter(breaks = c(1:5,6), palette = "Reds") + 
 # Add the fire boundary
 geom_sf(data=fire, fill="transparent", color="gray30", lwd=0.45) +
 geom_sf(data=points, color=alpha("black",0.8), size=0.85, shape=18) +
 coord_sf() +
 labs(fill="Days Since Ignition") +
 theme_void() +
 theme(legend.position="none",
       plot.title = element_text(hjust = 0.5),
       plot.margin = unit(c(1,1,1,1),'lines')) +
 # guides(fill = guide_colorbar(title.position="top", title.hjust = 0.5,
 #                               barwidth = 12, barheight = 0.5,
 #                               ticks=FALSE, draw.llim=FALSE)) +
 ggspatial::annotation_scale(location = "br", width_hint = 0.1,
                             pad_x = unit(0.2,"in"), pad_y = unit(0.2,"in"),
                             line_width = 0.5, text_pad = unit(0.15,"cm"),
                             height = unit(0.15,"cm")) +
 ggspatial::annotation_north_arrow(location = "br", which_north = "true",
                                   pad_x = unit(0.3,"in"), pad_y= unit(0.4,"in"),
                                   width = unit(0.8,"cm"), height = unit(0.8,"cm"),
                                   style = north_arrow_fancy_orienteering)
p3
gc()

# Save the image
ggsave(p3, file = "../../figures/FigureX_MartinFire.png",
       dpi = 300, bg="transparent") # adjust dpi accordingly

```

Long Draw Fire, OR (2012) - the 2nd fastest fire

```{r}

# Grab the start and end DOY
evt <- fired.fast %>% 
 filter(id %in% c(82548)) %>%
 mutate(last_day = lubridate::yday(last_date))
start_doy = min(evt$ig_day) # first DOY
end_doy = max(evt$last_day) # end DOY

# Set the geometries
st_agr(mtbs) <- "constant"
st_agr(ztrax) <- "constant"

# MTBS footprint
fire <- mtbs %>%
  filter(Event_ID %in% c("OR4239211789420120708")) %>%
  st_transform(st_crs(lambert.prj))

# ZTRAX point locations
points <- ztrax %>% 
  filter(YearBuilt <= 2012) %>%
  st_intersection(., fire) %>%
  st_transform(st_crs(lambert.prj))

# Convert to SpatVects
fire <- vect(fire)
points <- vect(points)

# Load the burndate raster (MCD64A1)
# Reproject the grid and crop to the fire perimeter
fire_dir <- paste0(maindir,"/fast-fires/data/spatial/mod/mcd64a1/")
mcd64 <- terra::rast('../../data/spatial/mod/mcd64a1/mcd64a1_annual_burndate_y2012.tif')
mcd64 <- terra::project(mcd64,crs(lambert.prj, proj=TRUE))
mcd64 <- terra::crop(mcd64,fire,mask=T)

# Convert to data frame
p4 <- mcd64 %>% as.data.frame(xy = TRUE) %>% mutate(
 value = as.integer(mcd64a1_annual_burndate_y2012)
) %>% 
 dplyr::select(-mcd64a1_annual_burndate_y2012) %>%
 mutate(value = if_else(value < start_doy, NA, value),
        value = if_else(value > end_doy, NA, value),
        value = value - start_doy,
        value = if_else(value > 5, 6, value)) %>%
 filter(value > 0) %>%
 drop_na() %>%
 ggplot() +
 geom_raster(aes(x = x, y = y, fill = value)) +
 # scale_fill_brewer(palette="Reds", na.value="white", direction=-1) +
 scale_fill_fermenter(breaks = c(1:5,6), palette = "Reds") + 
 # Add the fire boundary
 geom_sf(data=fire, fill="transparent", color="gray30", lwd=0.45) +
 geom_sf(data=points, color=alpha("black",0.8), size=0.85, shape=18) +
 coord_sf() +
 labs(fill="Days Since Ignition",
      subtitle = "") +
 theme_void() +
 theme(legend.position="none",
       plot.title = element_text(hjust = 0.5),
       plot.margin = unit(c(1,1,1,1),'lines')) +
 # guides(fill = guide_colorbar(title.position="top", title.hjust = 0.5,
 #                               barwidth = 12, barheight = 0.5,
 #                               ticks=FALSE, draw.llim=FALSE)) +
 ggspatial::annotation_scale(location = "br", width_hint = 0.1,
                             pad_x = unit(0.2,"in"), pad_y = unit(0.2,"in"),
                             line_width = 0.5, text_pad = unit(0.15,"cm"),
                             height = unit(0.15,"cm")) +
 ggspatial::annotation_north_arrow(location = "br", which_north = "true",
                                   pad_x = unit(0.3,"in"), pad_y= unit(0.4,"in"),
                                   width = unit(0.8,"cm"), height = unit(0.8,"cm"),
                                   style = north_arrow_fancy_orienteering)
p4
gc()

# Save the image
ggsave(p4, file = "../../figures/FigureX_LongDrawFire.png",
       dpi = 300, bg="transparent") # adjust dpi accordingly

```


